---
title: "ADSC1000 Solutions to Assumption Violations"
format: pdf
editor: visual
---

### Data Description:

A data.frame with 935 observations on 17 variables:

-   wage: monthly earnings
-   hours: average weekly hours
-   IQ: IQ score
-   KWW: knowledge of world work score
-   educ: years of education
-   exper: years of work experience
-   tenure: years with current employer
-   age: age in years
-   married: =1 if married
-   black: =1 if black
-   south: =1 if live in south
-   urban: =1 if live in SMSA (standard metropolitan statistical area)
-   sibs: number of siblings
-   brthord: birth order
-   meduc: mother's education (years)
-   feduc: father's education (years)
-   lwage: natural log of wage

## Example 1 from slides

```{r}

wage2 <- read.csv("wage2.csv")
library(car)

# lm1 <- lm(wage~age+iq+educ+meduc+hours+tenure+
#               kww+feduc+exper, data = wage2)


## Diagnostics ##

# # linearity #
# 
# plot(lm1,1) 
# 
# # Normality #
# plot(lm1,2)
# shapiro.test(rstandard(lm1))
# 
# 
# # Constant Variance #
# plot(lm1,3)
# ncvTest(lm1)
# 
# 
# # Independence #
# 
# vif(lm1)
# durbinWatsonTest(lm1)

```

### Required function

```{r}

panel.hist <- function(x, ...)
{
    usr <- par("usr")
    par(usr = c(usr[1:2], 0, 1.5) )
    h <- hist(x, plot = FALSE)
    breaks <- h$breaks; nB <- length(breaks)
    y <- h$counts; y <- y/max(y)
    rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}

panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
    usr <- par("usr")
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits = digits)[1]
    txt <- paste0(prefix, txt)
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}



```

## Example 2 from slides

```{r}

library(dplyr)

# na.omit(wage2) %>% 
#   select(wage,age,iq,educ,meduc) %>% 
#   pairs(
#       lower.panel = panel.cor,
#                  diag.panel = panel.hist,
#                  upper.panel = panel.smooth)
# 
# 
# 
# na.omit(wage2) %>% 
#   select(wage,hours,tenure,kww,feduc,exper) %>% 
#   pairs(
#       lower.panel = panel.cor,
#                  diag.panel = panel.hist,
#                  upper.panel = panel.smooth)

```

## Example 3 from slides

```{r}

vif(lm1)


```

## Example 4 from slides

```{r}
library(MASS)

# bc <- boxcox(lm1)
# lambda <- bc$x[which.max(bc$y)]
# # To transform your data:
# transform.df <- transform(wage2, BoxWages =
# wage∧(lambda))
# 
# lmBox <- lm()

## 2 ##

# wt <- 1/lm(abs(lm1$residuals)~lm1$fitted.values)$fitted.values^2
# 
# wls.model <- lm(,
#                 weights = wt)
# 
# summary(wls.model)


```

### Outliers Example

```{r}

 Outliers <- read.csv("Outliers.csv")


 boxplot(Outliers)
 plot(Outliers)


ggplot(data=Outliers,mapping = aes(x = X1, y = Y1)) +
  geom_point()+
  labs(title = "#5: Scatterplot of Y1 by X1",
       x = "X1",
       y = "Y1")+
 geom_text(x=3, y=20, label=paste("Correlation: ",round(cor(Outliers$X1,Outliers$Y1),digits=3))) + geom_smooth(method='lm') #Add regression line to plot



ggplot(data=Outliers,mapping = aes(x = X2, y = Y2)) +
  geom_point()+
  labs(title = "#5: Scatterplot of Y2 by X2",
       x = "X1",
       y = "Y1")+
 geom_text(x=3, y=20, label=paste("Correlation: ",round(cor(Outliers$X2,Outliers$Y2),digits=3))) + geom_smooth(method='lm') #Add regression line to plot



## Regression Outliers ##

lm5.1 <- lm(Y1 ~ X1 + X2, data = Outliers)
summary(lm5.1)

plot(lm5.1,5)

# Remove Outlier #
lm5.2 <- lm(Y1 ~ X1 + X2, data = Outliers[1:39,])
summary(lm5.2)

plot(lm5.2,5)

```

## Example 5 from slides

```{r}


# # Normality #
# plot(lm1,2)
# shapiro.test(rstandard(lm1))

## Outliers ##

# plot(lm1,3)
# plot(lm1,4)

# data[-c(1,4,6),]

```

## Example 6 from slides

```{r}



```
